Learning Network Size while Training with ShrinkNets

### 2018-02-09 paper/ICML/evaluation.tex

@@

Previous research~\cite{Scardapane2017} that attempts to optimize network size during training
 focused on simple datasets and simple (i.e., fully connected)
architectures. We believe that self-resizing
network are most important for problems where the architecture 
involves many layers, such that that it is impractical to explore the space of all possible
layer-size configurations. For that reason, we picked \texttt{CIFAR10} \cite{Krizhevsky2009}. It is an
image classification dataset containing $60000$ color images $(3 \times 32
\times 32)$, belonging to $10$ different classes, running on  the \texttt{VGG16} network \cite{Srivastava2014}. VGG16
consists of alternating convolutional layers and \textit{MaxPool} layers
interleaved by \textit{BatchNorm} \cite{DBLP:journals/corr/IoffeS15} and
\textit{ReLU} \cite{Nair2010}  layers. The two last layers are fully connected
layers separated by a \textit{ReLU} activation function.

>>>>

\ra{add here the impl. thingy from s3, so we give the context of how we use
shrinknets}

@@

\ra{this is related work}Previous research~\cite{Scardapane2017} that attempts to optimize network size during training
 focused on simple datasets and simple (i.e., fully connected)
architectures. We believe that self-resizing
network are most important for problems where the architecture 
involves many layers, such that that it is impractical to explore the space of all possible
layer-size configurations.\ra{until here} For that reason, we picked \texttt{CIFAR10} \cite{Krizhevsky2009}. It is an
image classification dataset containing $60000$ color images $(3 \times 32
\times 32)$, belonging to $10$ different classes, running on  the \texttt{VGG16} network \cite{Srivastava2014}. VGG16
consists of alternating convolutional layers and \textit{MaxPool} layers
interleaved by \textit{BatchNorm} \cite{DBLP:journals/corr/IoffeS15} and
\textit{ReLU} \cite{Nair2010}  layers. The two last layers are fully connected
layers separated by a \textit{ReLU} activation function.

### 2018-02-09 paper/ICML/impl.tex

\subsubsection{Neuron killing}
Once Switch layers are placed in a network and
initialized (sampled from the $\mathcal{N}(0, 1)$ distribution),

\textbf{Threshold strategy}:  We kill neurons based on a threshold (in absolute
value), this is the method used by deep compression. It is bad because it is
scale dependant and is not robust to noise in the gradients (explain that)
\\ \textbf{Sign change strategy}:  We look at when the sign change. This is also
sensitive to noise but has the advantage
\textbf{Sign variance strategy}: We measure the variance of the sign ($-1, 1$) of each component in the \textit{Switch Layer}, and consider it dead when it goes over a predefined threshold. \gl{Maybe we should add some intuition on why it makes sense}. \gl{We should probably explain the impact of $\gamma$ and report numbers from the expertiment that show that it does not hurt too much to kill neurons, we have the data in this experiment in the evaluation that we probably want to remove later.}

\subsection{Implementation}

We have implemented \textbf{Switch Layers} and the associated training procedure
as a library in pytorch~\cite{github}.
The layer can be freely mixed with other popular layers such as convolutional
layers, batchnorm layers, fully connected layers etc.
There are two key challenges that must be overcome while implementing 
\textbf{Switch Layers}.
First, since ShrinkNets kill off a large fraction of neurons, we must remove the
overhead of carrying around dead neurons.
And second, we must update the various network layers and training machinery (e.g.
optimizer state) to reflect the deactivation of neurons.
We address these challenges by augmenting every layer with additional information
about neuron status and implementing a \emph{neural garbage collection} 
mechanism which prunes deactivated neurons on-the-fly and updates the state of 
the training machinery.
% We do so by augmenting every layer with additional information such as what
% neurons have been deactivated.
% The layer then updates its size accordingly and this change is then propagated
% to other parts of the training machinery including the optimizer which must
% know the sizes of parameters.

A second implementation challenge we address with ShrinkNets is the potentially
higher inference time.
This increase in inference time is both due to the extra computation introduced 
by switch layers and because our switch layers are not currently optimized for 
CUDA.
Therefore, for now, our library provides a post-training routine to ``fold''
switch layers into existing, optimized layers.
Specifically, for every switch layer, we check if the parent or child of this 
layer is a linearly scalable layer (e.g. convolutional layer, fully connected layer,
batchnormalization layer) and if so, we fold the switch layer into this 
neighboring layer by multiplying through by the switch weights.
In the unlikely case that a switch layer is sandwiched between two non-linearly
scalable layers, we leave the switch layer as is.

% \subsection{Speeding up training}

% In order to have a practical implementation of ShrinkNets it is absolutely
% necessary to prune useless neurons as quick as possible. In order to try more
% different strategies we split this task in two subproblems: detecting
% potentially non-crucial neurons and setting their corresponding entry in the
% switch layer to exactly 0, and neural garbage collection that

% In our library we implemented a variant of the second strategy and the last one.

% \subsubsection{Neural garbage collection}

% It is possible to reduce the overhead of the training process by removing
% neurons as soon as they become deactivated by $\theta$ going to 0.
%If
%disabled neurons are not quickly removed, the overhead might cause the training
%process to be significantly slower than classic neural networks. 
% To do this, we implemented a \emph{neural garbage collection} mechanism which
% prunes deactivated neurons on-the-fly,  reducing the processing time
% and memory overhead. 
% To support this feature, it is crucial to understand the
% information flow between neurons and layers in the neural network. 
% We achieve
% this by representing such information flow as a graph. 
% Vertices represent layers,
% and edges are event-hubs responsible for propagating information about disabled
% neurons to the relevant layers. 
% As soon as a layer receive an event, it updates
% his parameters to reflect the new size. 
% Each operation is stored in a log that
% is used later to update other component of the system. 
% For example some
% optimizers like Adam [Ref] store state about each parameter, their state needs
% to be updated the exact same way and at the exact same time as layers in order
% to obtain apropriate parameter updates.

% \subsubsection{Speeding up Inference}
% \mpv{This may be moved to discussion or ``other considerations'' or impl details.}

% our post-processing code searches the 
% network graph to identify a parent or child layer the linear layer closest (child or parent) to every 
% switch layer and multiplies the weights of that layer by the corresponding
% switch layer weights.

% Our proposed \textit{Switch Layer} are relevant only during training, and can be
% removed from the network architecture during inference if they can be ``folded''
% into the nearest linear layer.
% Specifically, we provide a post-processing routine that will, for each switch 
% layer, search the network graph

% The sole puropose of \textit{Switch Layer} are to detect which neurons should be
% killed and which should be left. When the training is over we will never remove
% more neurons. Therefore, they loose all their interest at inference time. 
% To
% improve size and inference time we propose the following method to get rid of
% them without changing the output of the model (modulo floating point errors).
% The basic idea is to start from every \textit{Switch Layer} and try to find the
% nearset linear operator (child or parent) and multiply its weights by the values
% in the switch layer. For example we can merge Filter layers with surrounding
% Convolutional layer, Linear layer or even BatchNormalization [ref] if they are
% before the \textit{Switch Layer}. However we need to be extremely careful not to
% cross a non-linearity otherwise it would change the output of network.

>>>>

In this section we discuss practical aspects of ShrinkNets.

%The previous section described the switch layers, which deactivate neurons
%on-the-fly. To benefit from ShrinkNets in practice, we must overcome two
%challenges. First, to effectively reduce the network size we must remove the
%deactivated neurons. Since removed neurons will no longer contribute to
%improving prediction accuracy, we must be confident when removing them. Second,
%for inference, we must remove the switch layers from the network. We describe
%our solution for both challenges in this section.

\subsection{On-The-Fly Neuron Removal}
\label{neuron_killing}

Switch layers are initialized with weights sampled from $\mathcal{N}(0,1)$, and
their values change as part of the training process so as to swith \emph{on} or
{off} neurons. When neurons are deactivated by switch layers, we want to remove
them to shrink the network size. Because after removing a neuron it will no
longer contribute to improving prediction accuracy, we must devise strategies
that remove neurons while being robust to changing values. We observe three
strategies:

\textbf{Threshold strategy}: Under this strategy neurons are removed based on a
fixed threshold expressed in terms of its absolute value. This is the method
used by deep compression \cite{}. We found this method is not robust to noise in
the gradients---which occurs commonly when training networks with stochastic
gradient descent---because the threshold depends on the scale of each weight.

\textbf{Sign change strategy}: Under this strategy neurons are removed when the
weight value changes its sign. This strategy works well in practice but it is
also sensible to gradient noise. \ra{explain what's the consequence of that} If
we sample a gradient that is not representative of the dataset when performing
gradient descent then we might wrongly kill a neuron.\ra{last sentence is
unclear}

\textbf{Sign variance strategy}: We measure the exponential moving variance of
the sign ($-1, 1$) of each parameter in the \textit{switch layer}. When the
value of the exponential moving variances goes over a predefined threshold, we
consider the parameter is deactivating the neuron, and therefore we remove it.
This strategy introduces two extra parameters, one to control the behavior of
the inertia of the variance, and the threshold, but we have found it is the
best performing in practice.

The last two strategies perform the best in practice; they reduce the network
parameters by an order of magnitude with only a minimal impact on the final
error obtained by the model.

%According to our experiments on linear and logistic regressions the
%two last strategies reduce the number of parameter update performed during
%training by at least an order of magnitude with only minimal impact on the final
%error obtained by the model.

\subsection{Additional Optimizations for ShrinkNets}

\noindent\textbf{Preparing for Inference: } With ShrinkNets we obtain
reduced-sized networks while training, which is the first benefit towards faster
inference. This networks are readily available for inference. However, because
they include switch layers---and therefore more parameters---they introduce
unnecessary overhead at inference time. To avoid this overhead, we reduce the
network parameters by combining each switch layer with its respective network
layer by multiplying the respective parameters.  \ra{can't we just simply
prevent the following problem at network design time?}In the unlikely case that a switch
layer is sandwiched between two non-linearly scalable layers, we leave the
switch layer as is.

\noindent\textbf{Neural Garbage Collection: }ShrinkNets decide on-the-fly which
neurons to remove. Since ShrinkNets remove a large fraction of neurons, we must
dynamically resize the network at runtime to not carry on the 
unnecessary overhead. We implement a neural garbage collection method as part of
our library which takes care of updating the necessary network layers as well as
updating optimizer state to reflect the neuron removal.

%\ra{move to eval:implementation, retrieve the neuran garbace collection and
%place it somewhere here}
%We have implemented \textbf{Switch Layers} and the associated training procedure
%as a library in pytorch~\cite{paszke2017automatic}.  The layer can be freely
%mixed with other popular layers such as convolutional layers, batchnorm layers,
%fully connected layers etc.  There are two key challenges that must be overcome
%while implementing \textbf{Switch Layers}.  
%First, since ShrinkNets kill off a
%large fraction of neurons, we must remove the overhead of carrying around dead
%neurons.  And second, we must update the various network layers and training
%machinery (e.g.  optimizer state) to reflect the deactivation of neurons.  We
%address these challenges by augmenting every layer with additional information
%about neuron status and implementing a \emph{neural garbage collection}
%mechanism which prunes deactivated neurons on-the-fly and updates the state of
%the training machinery.

%A second implementation challenge we address with ShrinkNets is the potentially
%higher inference time.  This increase in inference time is both due to the extra
%computation introduced by switch layers and because our switch layers are not
%currently optimized for CUDA.  Therefore, for now, our library provides a
%post-training routine to ``fold'' switch layers into existing, optimized layers.
%Specifically, for every switch layer, we check if the parent or child of this
%layer is a linearly scalable layer (e.g. convolutional layer, fully connected
%layer, batchnormalization layer) and if so, we fold the switch layer into this
%neighboring layer by multiplying through by the switch weights.  
%In the unlikely
%case that a switch layer is sandwiched between two non-linearly scalable layers,
%we leave the switch layer as is.
